Hybrid Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-08-13 manuscript/main.tex
operator definition, except for the last

	\section{Data Objects and Model Formulation} \jongmin{still revising}
	For $k = 1, \ldots, K$, 
	we assume that the functional predictors $X_{ik}(t)$ are observed on the domain $0 \leq t \leq 1$, and belong to $\mathcal{F}$,
	which is defined as the space of real-valued functions on $[0,1]$
	which are square-integrable with respect to the Lebesgue measure $dt$ on $[0,1]$,
	and possess second  derivatives.
	Additionally, let $\mathbf{Z}=(Z_1, \ldots, Z_p)^\top$
	denote a $p$-dimensional multivariate scalar data in $\mathbb{R}^p$, assumed to be a random vector with finite first two moments, equipped with the Euclidean inner product and norm.
	% i.e., for $\mathbf{v_1}=(v_{11}, \ldots, v_{1p})^\top$ and $\mathbf{v_2}=(v_{21}, \ldots, v_{2p})^\top$ in $\mathbb{R}^p$, $\langle \mathbf{v_1}, \mathbf{v_2} \rangle = \mathbf{v_1}^\top \mathbf{v_2} = \sum_{r=1}^p v_{1r}v_{2r}$, and $\Vert \mathbf{v}_1 \Vert = \langle \mathbf{v}_1, \mathbf{v}_1 \rangle^{1/2} = (\sum_{r=1}^p v_{1r}^2)^{1/2}$.
	The hybrid random predictor and corresponding inner product is formally defined as follows.
	\begin{definition}[Hybrid predictor and inner product]\label{def:hilbert_space}
		We define a  hybrid predictor  $W_i$ as an ordered pair combining the functional predictors $X_{i1}(t), \ldots, X_{iK}(t)$ and a scalar predictor $\mathbf{Z}_i$:
		\begin{equation}\label{def:hybid_predictor}
			W_i := (X_{i1}, \ldots, X_{iK}, \mathbf{Z}_i)
			\in \mathcal{H} :=  \mathcal{F}^{\otimes K} \times \mathbb{R}^p
		\end{equation}
		The inner product between two hybrid predictors,
		$W_1 = (X_{11}, \ldots, X_{1K}, \mathbf{Z}_1)$
		and
		$W_2 = (X_{21}, \ldots, X_{2K}, \mathbf{Z}_2)$, is then defined as:
		\begin{equation}		\label{eq: hybrid inner product}
			\langle W_1, W_2\rangle_{\mathcal{H}}
			:=
			\sum \limits_{k=1}^K \int_0^1 X_{1k}(t) X_{2k}(t) \, dt
			+ 
			\omega
			\mathbf{Z}_1^\top \mathbf{Z}_2,
		\end{equation}
		for $\omega > 0$.
		The corresponding norm $\Vert \cdot \Vert_\mathcal{H}$ is derived from this inner product as $\langle \cdot,\cdot \rangle_\mathcal{H}^{1/2}$.
	\end{definition}
	In \eqref{eq: hybrid inner product}, 
	$\omega$ is a positive weight that needs to be pre-specified or estimated. It is mainly used to take into account heterogeneity between functional and scalar parts in terms of measurement scale and/or amount of variation (see Section \ref{subsec: Data Preprocessing}).

@@

	\subsection{Tucker's Criterion}\jongmin{notation not yet changed}
	We derive that the PLS components obtained from the first step of the proposed algorithm satisfy the Tucker's Criterion \citep{Tucker1938} extended to our scalar-on-hybrid regression model setting \eqref{eq: Hybrid functional model}. We omit $l$ in the notations and first define the cross-covariance terms between the response and predictors (observed versions if $l=1$ and residualized versions if $l \ge 2$): 
	\begin{equation*}
		\begin{gathered}
			\sigma_{YX} = (\sigma_{YX}^{(1)}, \ldots, \sigma_{YX}^{(K)}) = (E(YX^{(1)}),\ldots, E(YX^{(K)})) = E(YX) \in \mathcal{F} \\
			\sigma_{YZ} = (\sigma_{YZ,1}, \ldots, \sigma_{YZ,p})^\top = (E(YZ_1),\ldots, E(YZ_p))^\top = E(Y\mathbf{Z}) \in \mathbb{R}^p \\
			\Sigma_{YW}=(\sigma_{YX}, \sigma_{YZ}) =  (E(YX), E(Y\mathbf{Z}))=E(Y\mathbf{W}) \in \mathcal{H}
		\end{gathered}
	\end{equation*}
	We also introduce two cross-covariance operators, $\mathcal{C}_{YW}=E(\mathbf{W} \otimes_\mathcal{H} Y): \mathcal{H} \rightarrow \mathbb{R}$ and $\mathcal{C}_{WY}=E(Y \otimes \mathbf{W}): \mathbb{R} \rightarrow \mathcal{H}$, which respectively map $h=(f, \mathbf{v}) \in \mathcal{H}$ to $\mathbb{R}$ and $d \in \mathbb{R}$ to $\mathcal{H}$ as follows:
	\begin{align*}
		& \mathcal{C}_{YW} \mathbf{h} = E\{\langle \mathbf{W}, \mathbf{h} \rangle_\mathcal{H} Y \} = \langle \Sigma_{YW}, \mathbf{h} \rangle_\mathcal{H} = \sum \limits_{k=1}^K \int_0^1 \sigma_{YX}^{(k)}(t_k) f^{(k)}(t_k) dt_k + E(Y \mathbf{Z}^\top) \mathbf{v} \\
		& \mathcal{C}_{WY} d = E (\langle Y, d \rangle \mathbf{W} ) = E(Y\mathbf{W})d =  \Sigma_{YW} \, d. 
	\end{align*}

@@

The next proposition states that the vectors of estimated PLS scores for different iteration numbers are mutually orthogonal.
	\begin{proposition}	\label{proposition: orthnormality of PLS scores}
		Recall from Lemma \ref{proposition:closed_form_orthgonalization} that   $\widehat{\boldsymbol{\rho}}^{[l]}$ denote the $n$-dimensional vector whose elements consist of the $l$-th estimated PLS scores ($l=1, \ldots, L$) of $n$ observations.
		The vectors $\widehat{\boldsymbol{\rho}}^{[1]}, \widehat{\boldsymbol{\rho}}^{[2]}, \ldots, \widehat{\boldsymbol{\rho}}^{[L]}$  are mutually orthogonal in the sense that
		\begin{equation*}
			\widehat{\boldsymbol{\rho}}^{[l_1] \top}  \widehat{\boldsymbol{\rho}}^{[l_2] \top}= 0 \quad \textnormal{for} \quad \; l_1,l_2 \in \{1, \ldots, L\}, \; l_1 \ne l_2. 
		\end{equation*}
	\end{proposition}
	The proof of Proposition \ref{proposition: orthnormality of PLS scores} is provided in Appendix \ref{section:proof:proposition: orthnormality of PLS scores}.

@@

	\section{Simulations} \label{Simulations}

>>>>

	\section{Data Objects and Model Formulation} \jongmin{still revising}
	For $k = 1, \ldots, K$, 
	we assume that the functional predictors $X_{ik}(t)$ are observed on the domain $0 \leq t \leq 1$, and belong to $\mathcal{F}$,
	which is defined as the space of real-valued functions on $[0,1]$
	which are square-integrable with respect to the Lebesgue measure $dt$ on $[0,1]$,
	and possess second  derivatives.
	Additionally, let $\mathbf{Z}_i=(Z_{i1}, \ldots, Z_{ip})^\top$
	denote a $p$-dimensional multivariate scalar data in $\mathbb{R}^p$, assumed to be a random vector with finite first two moments, equipped with the Euclidean inner product and norm.
	% i.e., for $\mathbf{v_1}=(v_{11}, \ldots, v_{1p})^\top$ and $\mathbf{v_2}=(v_{21}, \ldots, v_{2p})^\top$ in $\mathbb{R}^p$, $\langle \mathbf{v_1}, \mathbf{v_2} \rangle = \mathbf{v_1}^\top \mathbf{v_2} = \sum_{r=1}^p v_{1r}v_{2r}$, and $\Vert \mathbf{v}_1 \Vert = \langle \mathbf{v}_1, \mathbf{v}_1 \rangle^{1/2} = (\sum_{r=1}^p v_{1r}^2)^{1/2}$.
	The hybrid random predictor and corresponding inner product is formally defined as follows.
	\begin{definition}[Hybrid predictor and inner product]\label{def:hilbert_space}
		We define a  hybrid predictor  $W_i$ as an ordered pair combining the functional predictors $X_{i1}(t), \ldots, X_{iK}(t)$ and a scalar predictor $\mathbf{Z}_i$:
		\begin{equation}\label{def:hybid_predictor}
			W_i := (X_{i1}, \ldots, X_{iK}, \mathbf{Z}_i)
			\in \mathcal{H} :=  \mathcal{F}^{\otimes K} \times \mathbb{R}^p
		\end{equation}
		The inner product between two hybrid predictors,
		$W_1 = (X_{11}, \ldots, X_{1K}, \mathbf{Z}_1)$
		and
		$W_2 = (X_{21}, \ldots, X_{2K}, \mathbf{Z}_2)$, is then defined as:
		\begin{equation}		\label{eq: hybrid inner product}
			\langle W_1, W_2\rangle_{\mathcal{H}}
			:=
			\sum \limits_{k=1}^K \int_0^1 X_{1k}(t) X_{2k}(t) \, dt
			+ 
			\omega
			\mathbf{Z}_1^\top \mathbf{Z}_2,
		\end{equation}
		for $\omega > 0$.
		The corresponding norm $\Vert \cdot \Vert_\mathcal{H}$ is derived from this inner product as $\langle \cdot,\cdot \rangle_\mathcal{H}^{1/2}$.
	\end{definition}
	In \eqref{eq: hybrid inner product}, 
	$\omega$ is a positive weight that needs to be pre-specified or estimated. It is mainly used to take into account heterogeneity between functional and scalar parts in terms of measurement scale and/or amount of variation (see Section \ref{subsec: Data Preprocessing}).

@@

	\subsection{Tucker's Criterion}\jongmin{notation not yet changed}
	We derive that the PLS components obtained from the first step of the proposed algorithm satisfy the Tucker's Criterion \citep{Tucker1938} extended to our scalar-on-hybrid regression model setting \eqref{eq: Hybrid functional model}. We omit $l$ in the notations and first define the cross-covariance term between the response and the functional predictors (observed versions if $l=1$ and residualized versions if $l \ge 2$): 
	\begin{equation*}
\sigma_{YX}
=
 (\sigma_{YX, 1} , \ldots, \sigma_{YX, K} )
  :=
\bigl(
\mathbb{E}
[Y_1 X_{11}]
,\ldots, 
\mathbb{E}
[Y_1  X_{1K}]
\bigr)   \in \mathcal{F}^{\otimes K},
	\end{equation*}
between the response and the scalar predictor:
\begin{equation*}
\sigma_{YZ} 
= 
(\sigma_{YZ,1}, \ldots, \sigma_{YZ,p})^\top 
:= 
\bigl(
\mathbb{E}[Y_1 Z_{11}]
,\ldots,
\mathbb{E}[Y_1 Z_{1p}]
\bigr)^\top %= E(Y\mathbf{Z})
\in \mathbb{R}^p.
\end{equation*}
Based on these definitions, we define the cross-covariance term between the response and the hybrid predictor as
\begin{equation*}
\Sigma_{YW}=(\sigma_{YX}, \sigma_{YZ}) :=  
\mathbb{E}
[ Y_1 W_1 ]
=
\bigl(
\sigma_{YX}, \sigma_{YZ}
\bigr) \in \mathcal{H}.
\end{equation*}
Based on these definitions, we introduce two cross-covariance operators.
First, $\mathcal{C}_{YW}= \mathbb{E}(W_1 \otimes_\mathcal{H} Y_1): \mathcal{H} \rightarrow \mathbb{R}$ 
maps  
  $h=(f_1, \ldots, f_K, \mathbf{v}) \in \mathcal{H}$ to a real number:
\begin{equation*}
\mathcal{C}_{YW}
\mathbf{h}
:=
\mathbb{E} 
\bigl[ 
\langle W_1, h \rangle_\mathcal{H} Y_1 
\bigr] 
= 
\langle \Sigma_{YW}, h \rangle_\mathcal{H}
=
\sum \limits_{k=1}^K \int_0^1 \sigma_{YX, k}(t) \, f_k(t) \, dt 
+ 
\sigma_{YW}^\top \mathbf{v}.
\end{equation*}
Next, $\mathcal{C}_{WY} = \mathbb{E}[Y \otimes \mathbf{W}]: \mathbb{R} \rightarrow \mathcal{H}$
maps  $d \in \mathbb{R}$ to a hybrid object in $\mathcal{H}$ as follows:
%%%%
\begin{align*}		
\mathcal{C}_{WY} d 
:= 
\mathbb{E} 
\bigl[
 \langle Y_1, d \rangle W_1 
\bigr] 
= 
\mathbb{E}
[ Y_1 W_1 d ] =  d \, \Sigma_{Y W}.
\end{align*}

@@

The next proposition states that the vectors of estimated PLS scores for different iteration numbers are mutually orthogonal.
	\begin{proposition}	\label{proposition: orthnormality of PLS scores}
		Recall from Lemma \ref{proposition:closed_form_orthgonalization} that   $\widehat{\boldsymbol{\rho}}^{[l]}$ denote the $n$-dimensional vector whose elements consist of the $l$-th estimated PLS scores ($l=1, \ldots, L$) of $n$ observations.
		The vectors $\widehat{\boldsymbol{\rho}}^{[1]}, \widehat{\boldsymbol{\rho}}^{[2]}, \ldots, \widehat{\boldsymbol{\rho}}^{[L]}$  are mutually orthogonal in the sense that
		\begin{equation*}
			\widehat{\boldsymbol{\rho}}^{[l_1] \top}  \widehat{\boldsymbol{\rho}}^{[l_2] }= 0 \quad \textnormal{for} \quad \; l_1,l_2 \in \{1, \ldots, L\}, \; l_1 \ne l_2. 
		\end{equation*}
	\end{proposition}
	The proof of Proposition \ref{proposition: orthnormality of PLS scores} is provided in Appendix \ref{section:proof:proposition: orthnormality of PLS scores}.

@@
